---
title: "MECH502_HW4_4"
author: "Jeremy Primus"
date: "September 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())           # clear variables
# Load the packages into R
library(knitr)
library(DEoptim)
```
# Intialization
```{r data input}
sigma = c(0, 141, 202, 252, 290, 319, 343, 360, 373, 390)
epsilon = c(0.00, 0.087, 0.172, 0.259, 0.339, 0.413, 0.482, 0.547, 0.608, 0.770)
plot(epsilon, sigma)
```

# Model selection

 Initial model selection:   
 Holloman
 
 Justification:   
   simplest model to fit (least parameters)        
                curve starts at the origin

# Model fitting
```{r fitting}
minFcn <- function(fitVars) {
  f <- sum((sigma - fitVars[1]*epsilon^fitVars[2])^2);
  return(f)
}
lower <- c(0,0)
upper <- c(1000,10)
set.seed(1234)                                  # set seed for reproducibility 
invisible(outDEoptim <- DEoptim(minFcn, lower, upper, DEoptim.control(itermax = 500, trace = FALSE)))     # run fit
summary(outDEoptim)
paramFits <- outDEoptim$optim$bestmem

# assign fit parameters
k <- paramFits[1] 
n <- paramFits[2]  

# identify where necking occurs
# necking criterion epsilon_t = n
sigma_neck = k*n^n

#plot(epsilon, 500*epsilon^0.5)
plot(epsilon, sigma, main = "True Stress-Strain") # plot true data
x <- seq(0, 0.8, by = 0.01)
lines(x, k*x^n)            # plot the fit
abline(v = n, col = "red") # mark where necking begins
```
 
  **Table 1.** Values of important parameters
```{r table of results, echo=FALSE}
df <- data.frame(format(c(k, n, sigma_neck), digits = 2), row.names = c("k", "n", "sigma_neck"))  # create a table
colnames(df) <- c("best fit value")
kable(df)
```
  
# Engineering stress-strain

```{r engineering stress-strain}
epsilon_e <- exp(epsilon)-1 # from 3.13
Ao_A <- exp(epsilon)        # from 3.25
sigma_e <- sigma*(1/Ao_A)   # from 3.14
plot(epsilon_e, sigma_e, main = "Engineering Stress-Strain")

```

# Discussion
  In the tradtional engineering stress-strain curve, the UTS defines an inflection point past which the material begins to neck and the strain is not uniform.  Since the strain is no longer hardening at a rate greater than geometric softening, we see a negative slope.  In the true stress-strain case, since the material is still hardening, the slope remains positive, although the slope may decrease, it will remain positive to failure.