---
title: "STAA575_HW3"
author: "Jeremy Primus"
date: "2/4/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Code and Results from homework 2
```{r Gibbs Sampler HW2, echo = FALSE}
rm(list=ls())           # clear variables
library(knitr)
library(coda)
library(mcmcplots)

# data collection
tis <- c(94.50, 15.70, 62.90, 126.0, 5.20, 31.40, 1.05, 1.05, 2.10, 10.50)     # durations of data collection
xis <- c(5, 1, 5, 14, 3, 19, 1, 1, 4, 22)     # number of failures
k <- length(xis)                              # number of parameters in the model (ie pumps)

# model parameters
alpha <- 1
c <- 1
d <- 1

iter <- 20000     # number of iterations 

# create iterable variables and initialize first values
b <- matrix(1,iter,1)
b[1] <- 2                                    # starting value of beta
theta <- matrix(1,iter,k)
theta[1,] <- xis/tis                         # data: failure rates 
predictedMean <- rep(0,k) 
compliant <- letters[1:k]

# Gibbs loop
for(t in seq(2, iter, length=iter)){
  b[t] <- rgamma(1, k*alpha+c, rate = sum(theta[t,])+d)
  for(i in seq(1, k, length=k)){
    theta[t,i] <-rgamma(1, xis[i]+alpha, rate = tis[i]+b[t])
  }
}

# identify 95% confidence interval
credibleIntervals <- HPDinterval(mcmc(theta), prob = 0.95)

# figure generation
for(j in seq(1, k, length=k)){
#  hist(theta[,j], main = paste('Pump ', j, ' Histogram'))
  predictedMean[j] <- mean(theta[,j])
  if (credibleIntervals[j,2] > 1){      # compliance criteria
    compliant[j] <- "no"}
  else {compliant[j] <- "yes"}
}
```
```{r summary table HW2, echo=FALSE}
# construct and present table
df = data.frame(predictedMean, credibleIntervals, compliant)
kable(df, 'markdown')

params <- matrix(c(rep(alpha, length(b)), b), nrow = length(b))
paramMeans <- rep(0,2)
paramMeans[1] <- alpha
paramMeans[2] <- mean(b)

paramCIIntervals <- HPDinterval(mcmc(params), prob = 0.95)
parametersdf = data.frame(paramMeans, paramCIIntervals, row.names = c("alpha", "beta"))
kable(parametersdf, 'markdown')

```
## Metropolis-within-Gibbs Sampler
```{r intialization}
rm(list=ls())           # clear variables
library(knitr)
library(coda)
library(mcmcplots)

# data collection
tis <- c(94.50, 15.70, 62.90, 126.0, 5.20, 31.40, 1.05, 1.05, 2.10, 10.50)     # durations of data collection
xis <- c(5, 1, 5, 14, 3, 19, 1, 1, 4, 22)     # number of failures
k <- length(xis)                              # number of parameters in the model (ie pumps)

# model parameters
c <- 1
d <- 1
mu <- 1

iter <- 10000     # number of iterations 

# create iterable variables and initialize starting values
alpha <- matrix(1,iter,1)
alpha[1] <- 1
beta <- matrix(1,iter,1)
beta[1] <- 1                                    # starting value of beta
theta <- matrix(1,iter,k)
theta[1,] <- xis/tis                            # data: failure rates 
predictedMean <- rep(0,k) 
compliant <- letters[1:k]

# initialize vectors for final table 
rownames <- letters[1:k]
paramMeans <-rep(0,2)
```
```{r Gibbs-MH algorithm}
# Gibbs-MH loop
for(t in seq(2, iter, length=iter)){
  # Metropolis-Hastings step for alpha (random walk)
  alphastar <- rexp(1, rate = 1/alpha[t-1])   # draw from proposal
  mHRatio <- (prod(theta[t-1,]^(alphastar-1))*beta[t-1]^(k*alphastar)*exp(-mu*alphastar)/gamma(alphastar))*dexp(alpha[t-1], rate = 1/alphastar)/(prod(theta[t-1,]^(alpha[t-1]-1))*beta[t-1]^(k*alpha[t-1])*exp(-mu*alpha[t-1])/gamma(alpha[t-1])*dexp(alphastar, rate = 1/alpha[t-1]))
  
  if (mHRatio >= 1){        # accept criteria 1        
      alpha[t] <- alphastar                    
    } else {
      u <- runif(1, 0, 1)     # Bernoulli draw
      if (u < mHRatio){       # accept criteria 2
        alpha[t] <- alphastar       
      } else {                # reject criteria
          alpha[t] <-alpha[t-1]       
      }}
      
  # Gibbs sampling
  beta[t] <- rgamma(1, k*alpha+c, rate = sum(theta[t,])+d) # conditional distribution for beta
  for(i in seq(1, k, length=k)){                           # conditional distribution for thetas
    theta[t,i] <-rgamma(1, xis[i]+alpha, rate = tis[i]+beta[t])
  }
}
```
```{r create plots and diagnostics}
# remove burn-in from results
theta <- theta[5001:10000,] 
alpha <- alpha[5001:10000,] 
beta <- beta[5001:10000,] 

# identify 95% confidence interval
credibleIntervals <- HPDinterval(mcmc(theta), prob = 0.95)

# histogram plots of theta distributions
for(j in seq(1, k, length=k)){
# doesn't request histogram  hist(theta[,j], main = paste('Pump ', j, ' Histogram'), xlab = paste('Theta ', j))
  predictedMean[j] <- mean(theta[,j])
  rownames[j] = paste('Pump', j)
  if (credibleIntervals[j,2] > 1){      # compliance criteria
    compliant[j] <- "no"}
  else {compliant[j] <- "yes"}
}
```
```{r final summary tables}
# construct and present table
thetadf = data.frame(predictedMean, credibleIntervals, compliant, row.names = rownames)
colnames(thetadf) <- c("Predicted Mean", "Lower Bound 95% CI", "Upper Bound 95% CI", "Compliant?")
kable(thetadf, 'markdown')

params <- matrix(c(alpha, beta), nrow = length(alpha))
paramMeans[1] <- mean(alpha)
paramMeans[2] <- mean(beta)

paramCIIntervals <- HPDinterval(mcmc(params), prob = 0.95)
parametersdf = data.frame(paramMeans, paramCIIntervals, row.names = c("alpha", "beta"))
kable(parametersdf, 'markdown')

```
## 2. Summary/Comparison
Implementing a prior onto alpha made the algorithm more complicated, and possibly more accurate, however, it did not change inferences made about pump compliance  

## 3. Caterplots


```{r caterplots}
# create a caterplot for the thetas
caterplot(theta)
abline(v=1)
title(main="Caterpillar plot of posterior distribution\n of pump failure rates",
xlab="# failures per 1000 hours",ylab="pump")
```
  
## 4. Sample Path and ACF

```{r sample path and acf}
# generate sample path plots
plot(theta[,1], type = 'n', main= "Sample Path: Theta 1")
lines(theta[,1])
acf(theta[,1])
plot(theta[,9], type = 'n', main= "Sample Path: Theta 9")
lines(theta[,9])
acf(theta[,9])
plot(alpha, type = 'n', main= "Sample Path: Alpha")
lines(alpha)
acf(alpha)
plot(beta, type = 'n', main= "Sample Path: Beta")
lines(beta)
acf(beta)
```
  
## Diagnostics Assessment  
The sample paths and ACFs do not show any rd flags indicating any major issues with the algorithm.  Sample paths show the parameter space is well explored, and the ACFs die off quickly.  There seems to be a slight oscillation visible in the theta values, but it is extremely small and I would assume that is just a negligible side effect of the algorithm.  It might explore the parameter space moving in one direction until it gets too far away, and then back the other way, and so on.  
