---
title: "STAA575_HW4-Parent"
author: "Jeremy Primus"
date: "2/11/2018"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())          # clear variables
library(knitr)
```
## Model Statement
```{r model definition}
# Model 1: original
model1.seeds.text="model {
   for (i in 1:N) {
      b[i]         ~ dnorm(0.0,tau)
      logit(p[i]) <- alpha0 + alpha1*x1[i] + alpha2*x2[i] +
                     alpha12*x1[i]*x2[i] + b[i]
      r[i]         ~ dbin(p[i],n[i])
   }
   alpha0  ~ dnorm(0.0,1.0E-6)  # intercept
   alpha1  ~ dnorm(0.0,1.0E-6)  # seed coeff
   alpha2  ~ dnorm(0.0,1.0E-6)  # extract coeff
   alpha12 ~ dnorm(0.0,1.0E-6)
   tau     ~ dgamma(1.0E-3,1.0E-3)    # 1/sigma^2
   sigma  <- 1.0/sqrt(tau)

}"

# Model 2: no interaction effect
model2.seeds.text="model {
   for (i in 1:N) {
      b[i]         ~ dnorm(0.0,tau)
      logit(p[i]) <- alpha0 + alpha1*x1[i] + alpha2*x2[i] + b[i]
      r[i]         ~ dbin(p[i],n[i])
   }
   alpha0  ~ dnorm(0.0,1.0E-6)  # intercept
   alpha1  ~ dnorm(0.0,1.0E-6)  # seed coeff
   alpha2  ~ dnorm(0.0,1.0E-6)  # extract coeff
   tau     ~ dgamma(1.0E-3,1.0E-3)    # 1/sigma^2
   sigma  <- 1.0/sqrt(tau)

}"

# Model 3: No root effect no interaction
model3.seeds.text="model {
   for (i in 1:N) {
      b[i]         ~ dnorm(0.0,tau)
      logit(p[i]) <- alpha0 + alpha1*x1[i] + b[i]
      r[i]         ~ dbin(p[i],n[i])
   }
   alpha0  ~ dnorm(0.0,1.0E-6)  # intercept
   alpha1  ~ dnorm(0.0,1.0E-6)  # seed coeff
   tau     ~ dgamma(1.0E-3,1.0E-3)    # 1/sigma^2
   sigma  <- 1.0/sqrt(tau)

}"

# Model 4: no random effect
model4.seeds.text="model {
   for (i in 1:N) {
      logit(p[i]) <- alpha0 + alpha1*x1[i] + alpha2*x2[i] +
                     alpha12*x1[i]*x2[i]
      r[i]         ~ dbin(p[i],n[i])
   }
   alpha0  ~ dnorm(0.0,1.0E-6)  # intercept
   alpha1  ~ dnorm(0.0,1.0E-6)  # seed coeff
   alpha2  ~ dnorm(0.0,1.0E-6)  # extract coeff
   alpha12 ~ dnorm(0.0,1.0E-6)
   tau     ~ dgamma(1.0E-3,1.0E-3)    # 1/sigma^2
   sigma  <- 1.0/sqrt(tau)

}"

model.seeds.list <- list(model1.seeds.text, model2.seeds.text, model3.seeds.text, model4.seeds.text)

# Model parameters
model.1.param <- c("alpha0", "alpha1","alpha2","alpha12","sigma")
model.2.param <- c("alpha0", "alpha1","alpha2","sigma")
model.3.param <- c("alpha0", "alpha1","sigma")
model.4.param <- c("alpha0", "alpha1","alpha2","alpha12")
model.params.list <- list(model.1.param, model.2.param, model.3.param, model.4.param)

```
## Data input
```{r data input}
#Data
  N <- 21
  r <-c(10, 23, 23, 26, 17, 5, 53, 55, 32, 46, 10, 8, 10, 8, 23, 0,  3, 22, 15, 32, 3)
  n <-c(39, 62, 81, 51, 39, 6, 74, 72, 51, 79, 13, 16, 30, 28, 45, 4, 12, 41, 30, 51, 7)
  x1 <-c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1)  # group 75 (0) or group 73 (1)
  x2 <-c(0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1)  # bean (0) or cucumber (1)
  seeds.data=list(N=N,r=r,n=n,x1=x1,x2=x2)
```
## Initial Values
```{r initial values}
#Initial Values:
#A diverse set of values for the GR statistic
# model 1 initse
inits1=list(
  list(tau =1,alpha0=0,alpha1=10,alpha2=-10,alpha12=5),
  list(tau =5,alpha0=5,alpha1=0,alpha2=-5,alpha12=10),
  list(tau =.1,alpha0=10,alpha1=5,alpha2=0,alpha12=-10))
# model 2 inits
inits2=list(
  list(tau =1,alpha0=0,alpha1=10,alpha2=-10),
  list(tau =5,alpha0=5,alpha1=0,alpha2=-5),
  list(tau =.1,alpha0=10,alpha1=5,alpha2=0))
# model 3 inits
inits3=list(
  list(tau =1,alpha0=0,alpha1=10),
  list(tau =5,alpha0=5,alpha1=0),
  list(tau =.1,alpha0=10,alpha1=5))
# model 4 inits
inits4=list(
  list(tau =1,alpha0=0,alpha1=10,alpha2=-10,alpha12=5),
  list(tau =5,alpha0=5,alpha1=0,alpha2=-5,alpha12=10),
  list(tau =.1,alpha0=10,alpha1=5,alpha2=0,alpha12=-10))

initslist <- list(inits1, inits2, inits3, inits4)
```
## Calls to model
```{r evaluate model performance, message= FALSE, warning= FALSE}
source("runEvaluateModels.R")
noOfModels <- length(model.seeds.list)
DIC <- list() 
plotstf = FALSE
for (ii in seq(1, noOfModels)){
  DIC[[ii]] <- runEvaluateModels(model.seeds.list[[ii]], model.params.list[[ii]], initslist[[ii]], seeds.data, plotstf)
  print(paste0("Model", ii))
  print(DIC[[ii]])
}
```
## Create a table of DICs

```{r create table }
modelMeanDICs <- c(102.4, 102.3, 99.45, 113.8)
modelPenalty <- c(10.99, 11.77, 16.01, 3.935)
modelPenDev <- c(113.4, 114.1, 115.5, 117.8)
df <- data.frame(modelMeanDICs, modelPenalty, modelPenDev, row.names = c("Model 1", "Model 2", "Model 3", "Model 4"))
colnames(df) <- c("DIC", "Effective No. of Parameters", "Penalized DIC")
kable(df)
```
  
## Interpret Results
     Model 1 has the second lowest DIC, and after assessing a penalty for the effective number of parameters, Model 1 has the lowest DIC.  Based upon this information, Model 1 will be selected for deeper analysis.  
```{r re run model 1}
plotstf = TRUE
runEvaluateModels(model.seeds.list[[1]], model.params.list[[1]], initslist[[1]], seeds.data, plotstf)
```
## Analysis (b-d)
The sample path and ACF plots indicated no obvious convergence issues with Model 1.  The posterior density suggests germination rates range from 33-55%, with a most probable germination rate of about 44%.  